Model Search in Relational Domains

### 2016-10-23 introduction.tex
intro and abstract

%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
%
Static program analysis based on abstract interpretation~\cite{CC77} has
been widely used to verify of properties of safety-critical systems. 
Standard static analyses~\cite{se2011} aim to compute program invariants as
fixed-points of abstract transformers.  Abstract states are chosen from a
lattice that has meet and join operations.  The meet precisely models set
intersection (or conjunction, taking a logical view), and the join
over-approximates set union (or disjunction).  The over-approximation in the
join operation is a key cause of loss of precision, which yields false
alarms.  Standard means to address false alarms in static analysis are
richer abstract domains or equipping the analysis with more precise set
unions -- which is often expensive since the analysis exhibits case
enumeration behaviour.

%%%%%%%%%%%%%%%%%%%%%%% ACDCL %%%%%%%%%%%%%%%%%%%%%
%
On the other hand, propositional SAT solvers are known to be highly
effective for solving complex propositional formulae, and are frequently
able to avoid enumerating the exponential number of cases.  The impressive
performance of modern solvers is credited to well-tuned decision heuristics
and clause learning algorithms in modern solvers.  Collectively, these
algorithms are referred to as \emph{Conflict Driven Clause Learning}
(CDCL)~\cite{cdcl}. The idea to lift these algorithms from the Boolean
domain to richer, arithmetic domains suggests itself.

>>>>

%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
%
Static program analysis based on abstract interpretation~\cite{CC77} has
been widely used to verify of properties of safety-critical systems~\cite{DBLP:conf/pldi/BlanchetCCFMMMR03}.
Standard static analyses~\cite{se2011} aim to compute program invariants as
fixed-points of abstract transformers.  Abstract states are chosen from a
lattice that has meet and join operations.  The meet precisely models set
intersection (or conjunction, taking a logical view), and the join
over-approximates set union (or disjunction).  The over-approximation in the
join operation is a key cause of loss of precision, which yields false
alarms.  Standard means to address false alarms in static analysis are
the use of richer abstract domains or equipping the analysis with more precise set
unions -- which is often expensive since the analysis exhibits case
enumeration behaviour.

%%%%%%%%%%%%%%%%%%%%%%% ACDCL %%%%%%%%%%%%%%%%%%%%%
%
On the other hand, propositional SAT solvers are known to be highly
effective for solving complex propositional formulae, and are frequently
able to avoid enumerating the exponential number of cases.  The impressive
performance of modern solvers is credited to well-tuned decision heuristics
and clause learning algorithms.  Collectively, these
algorithms are referred to as \emph{Conflict Driven Clause Learning}
(CDCL)~\cite{cdcl}. The idea to lift these algorithms from the Boolean
domain to richer, arithmetic domains suggests itself.

### 2016-10-23 paper.tex
double package
intro and abstract

%===============================================================================
\begin{abstract}
%===============================================================================
%
\emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a generalisation
of the well-known CDCL algorithm that is used in SAT solvers.
%
\Omit {  ACDCL alternates between a model search, which
performs over-approximate deduction with constraint propagation, and a
conflict analysis, which performs under-approximate abduction with heuristic
choice.
}
%
ACDCL has been successfully used with non-relational abstract domains such
as intervals to improve the performance of a floating-point decision
procedure.  Whereas the propositional reasoning used in SAT 
solvers can be rather easily lifted from the Boolean lattice to other 
non-relational domains, application to relational domains is 
challenging because of relational properties of deductions.
%
In this paper, we present the non-trivial liftings of the model search and
conflict analysis algorithms for relational abstract domains.  We have
implemented these algorithms inside the 2LS verification tool and
instantiated our analyser over a template polyhedra domain to check safety
properties of C programs.  We~evaluate the performance of our analyser in
comparison with program verifiers that use propositional-level solvers and
a static analysis tool.
%
\Omit {
on a set of benchmarks drawn from bit-vector regression 
in SV-COMP'16, bit-precise software models of hardware 
circuits auto-generated from v2c tool, and several 
bounds checking benchmarks 
}
We observe a reduction of two orders of magnitude in the number of 
decisions and conflict analyses compared to a SAT 
solver and higher precision compared to an abstract interpreter. 
\Omit{
along with stronger deductions and learnt clauses aided by the richer
abstract domains.
}
%
\end{abstract}

>>>>

%===============================================================================
\begin{abstract}
%===============================================================================
%
\emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a generalisation
of the well-known CDCL algorithm that is used in SAT solvers.
%
\Omit {  ACDCL alternates between a model search, which
performs over-approximate deduction with constraint propagation, and a
conflict analysis, which performs under-approximate abduction with heuristic
choice.
}
%
ACDCL has been successfully used with non-relational abstract domains such
as intervals to improve the performance of a floating-point decision
procedure.  Whereas the propositional reasoning used in SAT 
solvers can be rather easily lifted from the Boolean lattice to other 
non-relational domains, application to relational domains is 
challenging because of relational properties of deductions.
%
In this paper, we present the non-trivial liftings of the model search and
conflict analysis algorithms in SAT solver for relational abstract domains.  We have
implemented these algorithms inside the 2LS verification tool and
instantiated our analyser over a template polyhedra domain to check safety
properties of C programs.  We~evaluate the performance of our analyser in
comparison with program verifiers that use propositional-level solvers and
a static analyser based on abstract interpretation.
%
\Omit {
on a set of benchmarks drawn from bit-vector regression 
in SV-COMP'16, bit-precise software models of hardware 
circuits auto-generated from v2c tool, and several 
bounds checking benchmarks 
}
We observe a reduction of two orders of magnitude in the number of 
decisions and conflict analyses compared to a SAT 
solver and higher precision compared to an abstract interpreter. 
\Omit{
along with stronger deductions and learnt clauses aided by the richer
abstract domains.
}
%
\end{abstract}
